Implications of afforestation policies for global markets

### 2024-02-27 paper.Rmd
Update paper.Rmd

To understand the impact of potential regional and global afforestation actions, an important first step was creating a realistic baseline scenario that depicts the likely changes between over the decade under study: 2023-2033. This was done by taking into account key drivers of change like growth in population and GDP. Without accounting for this, results would likely have been biased as they would have failed to account for the increasing pressure to convert forests into agricultural land. Data for this were drawn from sources such as CEPII and the United Nations.

@@

Then, we examined what it would take given the pressures of a still growing global population with the concomitant increasing appetite for foodstuffs to end, on net, the practice of deforestation globally. Results suggest that it would require a tax of roughly 70.3 percent to be assessed on the value of the converted land for the deforestation rate to fall globally to roughly zero. The economic impact of such a policy measure would not be limited to the cost of the tax, it would also change impact food production and global trade. For example, with this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed. For the United States, agricultural land would be 2.1 percent less (Table 1). Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-8.0 percent), Brazil (-8.5 percent), and Russia (-9.3 percent).

>>>>

To understand the impact of potential regional and global afforestation actions, an important first step was creating a realistic baseline scenario that depicts the likely changes between over the decade under study: 2023-2033. Data for this were drawn from sources such as CEPII and the United Nations. Using this data, we were able to control for the impact of key drivers of change like growth in population and GDP. Without accounting for this, results would likely have been biased as they would have failed to account for the increasing pressure to convert forests into agricultural land.

@@

Then, we examined what it would take given the pressures of a still growing global population and an increasing appetite for foodstuffs for the practice of deforestation to end, on net, globally. Results suggest that it would require a tax of roughly 70.3 percent to be assessed on the value of the converted land for the deforestation rate to fall globally to roughly zero. The economic impact of such a policy measure would not be limited to the cost of the tax, it would also change impact food production and global trade. For example, with this policy in place, most regions of the world would see significant declines in the amount of agricultural land in 2033 relative to the baseline scenario where land conversion was not taxed. For the United States, agricultural land would be 2.1 percent less (Table 1). Larger decreases would happen in other major agricultural production regions such as China (-2.5 percent), South Asia (-8.0 percent), Brazil (-8.5 percent), and Russia (-9.3 percent).
